|

### 2021-07-29 prova_knit_finale.Rmd
corretto errore tabella
added cite

Usually, PF requires a resampling step at each stage of the algorithm. Recall that, in general, resampling determines a trade-off between the increase in the variance at the resampling stage and a decrease in the variance when future states are sampled from the transition kernel.\
"Dead" particles (i.e., those with small weights) are  likely to be removed when resampling takes place and past weights get ``reset" (usually, they are set to be equal to $1/N$). Otherwise, the \textit{curse of dimensionality} problem will emerge and the variance of the weights will diverge over time as the variance of past weights increases as new updates occur period-by-period. Figure ***** shows an example of a PF where SIS is used: note that the $ESS$ drops dramatically after few periods, meaning that the variance of the importance weights diverges. This motivates the use of a SIR structure for the PF algorithm.\ \
Lastly, it is now common practice to use a mixed approach with respect to resampling (recall that resampling has an additional computational cost). Generally, resampling is done \textit{adaptively}, when some condition is met. A standard criterion that is used (and that we adopt in our implementations of the PF) is to resample when the $ESS$ drops below some threshold $ESS_{min}\in [1,N]$:
\begin{equation}
    ESS(W_{t-1}^{1:N})<ESS_{min}
\end{equation}

@@

We are now ready to establish, informally, the optimality result about the transition kernel. Let us denote by  $G_t=[P_t(x_{t-1},dx_t)\cdot f_t(y_t|x_t)dy_t]/M_t(x_{t-1},dx_t)$ the \textit{incremental weights} (i.e., the updating factor) and by $M_t^*(x_{t-1},dx_t)$ the optimal kernel. It can be proved\footnote{For a formal proof see \cite{**}, th. 10.1 p. 142.} that the optimal kernel, i.e., the one that minimizes the variance of the incremental weights $G_t$ (and of the time $t$ weights themselves), is:

@@

It is possible to derive an optimality result for the auxiliary functions\footnote{Details in ***** th 10.2, p. 148}. In the special case in which the transition kernel used for sampling corresponds with the optimal one $M_t^*$, the optimal auxiliary function (i.e., the auxiliary function that minimizes the variance of the inferential weights) becomes:

@@

\item Normalize the auxiliary weights, for $n=1,...,N$:  $\tilde W_t^n:=\frac{\tilde w_t^n}{\sum_{m=1}^N\tilde w_t^m}$.\footnote{It is worth mentioning that the original APF included a final resampling step at the end of each iteration. However, our formulation is more efficient since having two resampling steps might introduce additional Monte Carlo variance. Se ***Johansen p. 120}

@@

@@

>>>>

Usually, PF requires a resampling step at each stage of the algorithm. Recall that, in general, resampling determines a trade-off between the increase in the variance at the resampling stage and a decrease in the variance when future states are sampled from the transition kernel.\
"Dead" particles (i.e., those with small weights) are  likely to be removed when resampling takes place and past weights get ``reset" (usually, they are set to be equal to $1/N$). Otherwise, the \textit{curse of dimensionality} problem will emerge and the variance of the weights will diverge over time as the variance of past weights increases as new updates occur period-by-period. Next, we show an example of a PF where SIS is used: note that the $ESS$ drops dramatically after few periods, meaning that the variance of the importance weights diverges. This motivates the use of a SIR structure for the PF algorithm.\ \
Lastly, it is now common practice to use a mixed approach with respect to resampling (recall that resampling has an additional computational cost). Generally, resampling is done \textit{adaptively}, when some condition is met. A standard criterion that is used (and that we adopt in our implementations of the PF) is to resample when the $ESS$ drops below some threshold $ESS_{min}\in [1,N]$:
\begin{equation}
    ESS(W_{t-1}^{1:N})<ESS_{min}
\end{equation}

@@

We are now ready to establish, informally, the optimality result about the transition kernel. Let us denote by  $G_t=[P_t(x_{t-1},dx_t)\cdot f_t(y_t|x_t)dy_t]/M_t(x_{t-1},dx_t)$ the \textit{incremental weights} (i.e., the updating factor) and by $M_t^*(x_{t-1},dx_t)$ the optimal kernel. It can be proved\footnote{For a formal proof see Chopin and Papaspiliopoulos (2020), th. 10.1 p. 142.} that the optimal kernel, i.e., the one that minimizes the variance of the incremental weights $G_t$ (and of the time $t$ weights themselves), is:

@@

It is possible to derive an optimality result for the auxiliary functions\footnote{Details in Chopin and Papaspiliopoulos (2020) th 10.2, p. 148}. In the special case in which the transition kernel used for sampling corresponds with the optimal one $M_t^*$, the optimal auxiliary function (i.e., the auxiliary function that minimizes the variance of the inferential weights) becomes:

@@

\item Normalize the auxiliary weights, for $n=1,...,N$:  $\tilde W_t^n:=\frac{\tilde w_t^n}{\sum_{m=1}^N\tilde w_t^m}$.\footnote{It is worth mentioning that the original APF included a final resampling step at the end of each iteration. However, our formulation is more efficient since having two resampling steps might introduce additional Monte Carlo variance. See Johansen and Evers (2007), p. 120}

@@

Johannes, M., Evers, L. (2007). Monte Carlo Methods. Lecture Notes. University of Bristol, Department of Mathematics.\

@@

Petris, G., Petrone, S., Campagnali, P. (2010). Dynamic Linear Models with R. Springer.\
